# Protection Island 1

## Questions

The information below is taken from the following source:
Newcomb, HR. 1940. [Ring-necked pheasant studies on Protection Island in the Strait of Juan de Fuca](https://ir.library.oregonstate.edu/concern/graduate_thesis_or_dissertations/js956j801?locale=en), Washington. MS thesis. Oregon State University.

a.	Pheasant chicks are born during the summer.
b.	In May 1937, 10 pheasants were introduced to the island. Before the next breeding season there were 35.
c.	November 10, 1938 a census estimated 110 pheasants.
d.	October 13, 1939 a census estimated 400 pheasants.

1. Read and complete all the exercises in Chapter 11 of Quantitative skills for biology [Link](https://ahurford.github.io/quantitative-training-guide/) 
2. All the examples of making a graph in Chapter 11 of [Quantitative skills for biology](https://ahurford.github.io/quantitative-training-guide/) have lists of _numbers_ as the values that are plotted on the axes. We would like to make a graph of the data listed in b.-d., however, we will face a problem since the x-axis values are dates, rather than numbers. We will consider two possible solutions:

i. Use a built-in `R` function, so that dates can be plotted on the x-axis; and

ii. Convert the dates to number of days since a reference date. Now the dates are numbers and these values can be plotted on the x-axis.

In this question, we will proceed with option i. The function we will use is `as.Date()`. You can learn how to use this function using an internet search or by typing the following into your `Console`:

```{r,eval=FALSE}
?as.Date
```

These files can be difficult to understand. A good way to proceed is to experiment with the function in the `Console`. Try these:

```{r,eval=FALSE}
as.Date(2012-01-31, format = %Y-%m-%d)
as.Date("2012-01-31", format = "%Y-%m-%d")
```

Note that only the second command is error-free. The first command fails needs to contain `""` because the date argument for the `as.Date()` function must be a character string (see `?character`). It is also possible to omit the format argument and just code `as.Date("2012-01-31")`. The help file notes that when the format argument is not specified, the help file states that formats will be tried one by one and an error will be returned if none work. It is advisable to specify the format, as allowing the function to infer the format could introduce errors.

Section 6.9 of [Quantitative skills for biology](https://ahurford.github.io/quantitative-training-guide/) describes how to make a vector. We need to make a vector of the dates so that we can make our plot. For example,

```{r,eval=FALSE}
x <- as.Date(c("2012-01-31", "2012-03-05", "2013-01-11"), format = "%Y-%m-%d")
```

Having completed Chapter 11 of [Quantitative skills for biology](https://ahurford.github.io/quantitative-training-guide/) and having learned how to work with dates, you should now be able to make plot using the information in b.-d. above. The graph you hand in needs to have descriptive axeses and a figure caption. You may put these elements together using a word processing software such as _Microsoft Word_.

3. Next, we try approach ii. to make the graph. Under approach ii. we will work with the dates by converting them to the number of days since a reference date. To do this we will use the `julian()` function, which is part of the `chron` package.

Read Section 4.4 of [Quantitative skills for biology](https://ahurford.github.io/quantitative-training-guide/) regarding installing packages. Install the package `chron` using either the `Install` button on the `Packages` tab, or by using the command `install.packages("chron")` in the `Console` window. Note that the package is only available for use once you check the box on the `Packages` tab or by running the following command in the `Console`:

```{r,eval=FALSE}
require("chron")
```

After the `chron` package is loaded we can then query the `julian` function,

```{r,eval=FALSE}
?julian
```

or use an internet search. As the help files can be difficult to understand, another approach to is to try out the function. Try the following:

```{r,eval=FALSE}
julian(1,1,1970)
julian(1,2,1970)
julian(2,1,1970)
julian(1,1,1971)
julian(1,1,1969)
```

Which argument position of `julian()` function corresponds to the month?


1. Read and complete all the exercises in Chapter 11 of:
Quantitative skills for biology [Link](https://ahurford.github.io/quantitative-training-guide/)

Graph the population size (y-axis) versus time (x-axis). You will need to convert the dates into days since a specific reference point.

1.	For the discrete time exponential (geometric) growth model, ideally the census points should be a fixed amount of time apart, i.e., a day, a week, or a year. Provide 1-3 sentences discussing the timing of the census points for Protection Island.

2. 

2.	
